Rapid and Accurate Global PTM Discovery (G-PTM-D) Using Post-Acquisition Spectral Calibration and Multi-Notch Searches

### 2017-07-12 paper.tex
ajdsfk

@@

@@

>>>>

In the original G-PTM-D manuscript, an open mass search (+/- 200-500Da) was used to create a histogram of mass difference peaks, from which a number of PTMs could be inferred.
A perl script was used to mine those mass differences to modify a proteomics database with potential PTMs.
This database was used for a final search and a number of PTMs were identified.
There were several limitations to that work.
One, the open mass search was really slow.
Two, uncalibrated data made it hard to discern closely spaced modifications.
Three, certain mass differences with high FDRs were incorrectly allowed.
Four, missed monoisotopic peaks and co-isolated peaks were not handled.
Five, separate programs were required (both Morpheus and a perl script).
MetaMorpheus resolves all of these issues and more.
Below we will show how each individual enhancement to G-PTM-D improved performance over the original work and then we will show at the end, the combined effect of all enhancements.

@@

[THERE IS SUPPOSED TO BE A COOL FIGURE HERE]

G-PTM-D originally added PTMs with masses that were included in the UniProt PTM list. 
We have since increased the variety of PTMs added but including a number of new notches in the G-PTM-D first pass search. The new notches correspond to metal adducts (e.g. Na and K), glycosylations., sample handling and processing artifacts (e.g. carbamylation, ammonia loss), lipidation, and etc.
MSFragger has limited mass range as reported. We identified peaks with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
Some of these problematic histogram peaks are illustrated in Figure, including peaks corresponding to masses of Sulfonation and Phosphorylation, and a problematic peak at mass of Leucine, with a high FDR.
Some Histogram Peaks For Mouse Data Some Histogram Peaks For Mouse Data We identified some specific mass shifts that are characterized by anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals .
MSFragger also doesn’t automatically calibrate spectra prior to searching.
Therefore, mass differences from closely spaced PTMs (e.g. phosphorylation and sulfonation) can easily be misinterpreted.
MetaMorpheus has built in functionality to identify multiple peptides from each tandem mass spectrum resulting in improved identification of peptides and peptide modifications.

@@

We observe that when using an incomplete database that does not include comprehensive modification and sequence variation information, a notch search with notches at monoisotopic peaks is superior to either of the single-notch searches.
